%!TEX root = ../thesis.tex
\chapter{Background}

Dynamic mapping is becoming an increasingly important requirement for digital musical instruments. This chapter surveys currently available tools that allow for manipulation of musical and non-musical networks in real time. The first section presents a review of mapping itself, both from a theoretical and a musical standpoint. This portion also introduces the libmapper application programming interface. The second section reviews relevant work in the visual representation of information. The third section of this chapter describes applicable techniques in user interface design. Finally, a review of user interfaces for mapping is presented.

\section{Mapping}

	At the most fundamental level, \emph{mapping} is the act of associating two or more sets of information. Mappings can be mathematical, computational, linguistic (like translation), geographic, or even poetic\footnote{What is metaphor if not the association of unlike things?}. Within the context of DMI design mapping is the relationship between sensor outputs and synthesis inputs. The entire character of a new instrument can be drastically altered though mapping, even while control surface and sound source are held constant \shortcite{hunt_mapping_is_important}. As a result, the theoretical formalism of mapping becomes yet another necessary tool in the modern instrument designer's arsenal.

	\subsection{Mapping Theory}
	\subsubsection{Types of Mappings}

	From the perspective of mathematics, the term \emph{mapping} is very nearly synonymous with \emph{function} \cite{native_set_theory}, as both describe how one set of numbers corresponds with another. The first group is commonly refered to as the \emph{domain} and the second as the \emph{codomain} or \emph{range}. An in-depth review of functions in mathematics is beyond the scope of this thesis, however a few fundamental examples will be useful for reference in section \ref{sec:mappingforDMIs}. The following are instances of two basic types of mathematical functions:

	\begin{equation} y = 2x - 1 \label{eq:one-to-one} \end{equation} 
	\begin{equation} y = x^2  \label{eq:many-to-one}  \end{equation}

	Each function takes a single input value (\emph{x}) and \emph{maps} that number onto its range (\emph{y}). 
	%For example, an input of \textbf{5} maps to \textbf{9} in equation \ref{eq:one-to-one}, while the same input results in an output \textbf{25} for equation \ref{eq:one-to-many}. 
	The fact that each of these equations take in only a single number as input, and output a single number in turn, means they can be graphed in a two dimensional space. This is not necessarily the case, as functions can input and output lists of numbers (vectors). Mathematically they are not very interesting, but they represent two fundamentally different \emph{kinds} of functions.

	\begin{figure}[h]
		\centering
		\begin{tikzpicture}
			\begin{axis}[my style, xtick={-2,-1,...,2}, ytick={-2,-1,...,2}, xmin=-2, xmax=2, ymin=-2, ymax=3]
				\addplot[domain=-100:100]{2*x-1}; 
			\end{axis}
		\end{tikzpicture}	
	\caption{The function described in equation \ref{eq:one-to-one}, graphed in two dimensions.}
	\label{fig:one-to-one_graph}
	\end{figure}


	For equation \ref{eq:one-to-one} each input value has \emph{one and only one} corresponding output value. The same is true if the function is to be inverted, as each output value corresponds to only one input value. The domain is simply a scaled and shifted version of the range. The mapping's \emph{one-to-one} nature can clearly be seen in figure \ref{fig:one-to-one_graph}.

	\begin{figure}[h]
		\centering
		\begin{tikzpicture}
			\begin{axis}[my style, xtick={-3,-2,...,3}, ytick={-3,-2,...,3}, xmin=-3, xmax=3, ymin=-1, ymax=4]
				\addplot[domain=-3:3]{x^2}; 
			\end{axis}
		\end{tikzpicture}
	\caption{Equation \ref{eq:many-to-one} projected on the Cartesian plane.}
	\label{fig:many-to-one_graph}
	\end{figure}


	This is not the case for equation \ref{eq:many-to-one}, for although each input has only one output, single positions in the codomain can have multiple corresponding inputs (e.g. both $3^2$ \emph{and} $-3^2$ are equal to 9). Thus we can consider equation \ref{eq:many-to-one} to be a canonical example of a \emph{many-to-one} mapping. In figure \ref{fig:many-to-one_graph} the range of the function is wrapped back onto itself such that a horizontal line could intersect the function twice.

	Two more categories of mapping are relevant to instrument design, an example of each:  

	\begin{equation} y = \pm\sqrt{x} \label{eq:one-to-many} \end{equation} 
	\begin{equation} y = \pm\sqrt{1 - x^2} \label{eq:many-to-many} \end{equation} 

	They are not considered to be functions by mathematicians, but are nonetheless important for our purposes. In equation \ref{eq:one-to-many} a single input can result in multiple outputs (an input of 4 results in the output of \emph{both} 2 and -2), yet each output has only a single input. This is simply the inverse function of equation \ref{eq:many-to-one}, and is called a \emph{one-to-many} mapping. On a graph of such a mapping, a \emph{vertical} line may cross at multiple points. The final equation is that of a circle centered at the origin with a radius of one. This is an example of a \emph{many-to-many} mapping, as both it an its inverse result in multiple outputs from a single input.

	\begin{figure}[h]
		\centering
		\begin{tikzpicture}
			\begin{axis}[my style, xtick={-1,...,1}, ytick={-1,...,1}, xmin=-1.5, xmax=1.5, ymin=-1.5, ymax=1.5]
				\addplot[domain=-1:1]{sqrt(1 - x^2)}; 
				\addplot[domain=-1:1]{-sqrt(1 - x^2)}; 
			\end{axis}
		\end{tikzpicture}
	\caption{Equation \ref{eq:many-to-many}, a many-to-many mapping.}
	\end{figure}

	Though a graphical plane is the most common way for mathematicians to visualize two-dimensional functions, drawing the direct association between input and output will be more useful going forward.

	\begin{itemize}
		\item continuous vs discreet
	\end{itemize} 

	\subsection{Mapping for Digital Musical Instruments} \label{sec:mappingforDMIs}
		\begin{itemize}
			\item Cite vijay about the inputs-outputs/outputs->inputs/sources->destinations thing
			\item many2one/one2one/one2many/many2many, certainly can cite a gestures class readgin \shortcite{wanderley}
			\item These two contexts have also been referred to as a systems point of view and a functional point of view respectively [11].  Systems = network, functional = connection properties Two types of mapping \cite{two_types_of_mapping}
		\end{itemize}
	\subsection{libmapper}
		joe's libmapper paper: \shortcite{malloch}
		joe's other paper? (earlier), his master's thesis
		\subsubsection{Open Sound Control}
		OSC: \shortcite{osc}
		\subsubsection{Structure of libmapper Networks}
		\subsubsection{Control of libmapper Devices and Signals}
	\begin{enumerate}
		\item GDIF: \shortcite{GDIF} Describes the namespace that libmapper uses, ``Ideally, it should be possible to store all sorts of data from various commercial and custom made controllers, motion capture and computer vision systems, as well as results from differ- ent types of gesture analysis, in a coherent and consistent way. This would make it possible to use the information with different software, platforms and devices, and also allow for sharing data between research institutions.'' 
		\item disembodied performance
		\item Wanderley's mapping paper \shortcite{wanderley} 
		\item Jamoma \shortcite{jamoma}
		\item surely some other stuff from class
	\end{enumerate}

\section{Data Visualization}
	\subsection{Graphical Perception}
		\subsubsection{Heirarchical Structures}
		\subsubsection{Dense Information}
	\subsection{Visualization Techniques}
		\subsubsection{Filtering}
		\subsubsection{Spark Lines}
		\subsubsection{Dash Plots}
	\subsection{Visualization Systems}
		Allosphere?, Braun Braun: view OSC data flows \shortcite{braun}, HEB?
	\begin{enumerate}
		\item Allosphere? :\shortcite{allosphere}
		\item Heirarchical edge bundling: \shortcite{HEB}
		\item Tukey: \shortcite{tuckey}
		\item Envisioning information: \shortcite{tufte1}
		\item Beautiful Evidence: \shortcite{tufte2}
		\item The other Tufte book I have at home.
		\item OSC data flows with Braun \shortcite{braun}
	\end{enumerate}

\section{User Interface Design}
	\subsection{A Brief History of Electronic User Interfaces}
	\subsection{Task Analysis}
	\subsection{Recall and Recognition?}
	\subsection{Collaborative Network Interfaces}
		MPG Care Package \shortcite{MPGcarepackage}
	\subsection{The Model-View-Controller Architecture}
		MVC Krasner Pope \shortcite{MVC_krasnerpope}
	\subsection{User Centric Design}
		Organizational context \shortcite{usd}
		Usability testing \shortcite{usd_corry}
		Information professionals \shortcite{usd_schulze}
	\begin{enumerate}
		\item Inclusive interconnections \shortcite{inclusiveinterconnections}
		\item Sense Stage \shortcite{senseStage}
	\end{enumerate}

\section{Relevant User Interfaces}
	\subsection{Junxion}
		Junxion \shortcite{junxion}
	\subsection{Osculator}
		Osculator: mapping OSC stuff \shortcite{osculator}
	\subsection{Other Similar Interfaces}
		Integra \shortcite{integra}
		Eaganmatrix: GRID VIEW! \shortcite{eaganmatrix}
		Patchage: a linking, dragging, connecting interface \shortcite{patchage}
	\subsection{Prior Interfaces for libmapper} \label{sec:priorGUIs}
		Vizmapper \shortcite{vizmapper}
	

\section{Summary}
	