\documentclass [12pt,letterpaper]{report}

% Standard packages
\usepackage{amsmath}		% Extra math definitions
\usepackage{graphics}		% PostScript figures
\usepackage{setspace}		% 1.5 spacing
\usepackage{longtable}          	% Tables spanning pages
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[ampersand]{easylist}

% Aaron's Packages
\usepackage{Style/ichago}   % The ichago references style
\usepackage{url}            % For formatting urls
\usepackage{listings}       % For formatting code
\usepackage{color}
\usepackage{sidecap}
\usepackage{wrapfig}
\usepackage{pgfplots}

\pgfplotsset{my style/.append style={axis x line=middle, axis y line=middle, xlabel={$x$}, ylabel={$y$}, axis equal }}

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}

\lstset{frame=tb,
  language=C,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=none,
  numberstyle=\tiny\color{gray},
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  breaklines=false,
  breakatwhitespace=false,
  frame = single,
  lineskip = 0pt,
  tabsize=3
}

% Custom packages
\usepackage[first]{Style/datestamp}			% Datestamp on first page of each chapter
\usepackage[fancyhdr]{Style/McECEThesis}	% Thesis style
\usepackage{Style/McGillLogo}			% McGill University crest

% $Id: ThesisEx.tex,v 1.1 2005/06/09 12:48:46 kabal Exp $

\usepackage{color}
\def\headrulehook{\color{red}}		% Color the header rule

%===== page layout
% Define the side margins for a right-side page
\insidemargin = 1.3in
\outsidemargin = 0.8in

% Above margin is space above the header
% Below margin is space below footer
\abovemargin = 1.1in
\belowmargin = 0.75in

%========= Document start

\begin {document}

%===== Title page

\title{A Flexible Tool for the Visualization and Manipulation of Musical Mapping Networks}
\author{Aaron Henry Krajeski}
\date{\Month\ \number\year}
\organization{%
    \\[0.2in]
    \McGillCrest {!}{1in}\\	% McGill University crest
    \\[0.1in]
    Department of Music Technology\\
    Schulich School of Music, McGill University\\
    Montreal, Canada}
\note{%
    {\color{red} \hrule height 0.4ex}
    \vskip 3ex
    A thesis submitted to McGill University in partial fulfillment of the
    requirements for the degree of Master of Arts.
    \vskip 3ex
    \copyright\ \the\year\ Aaron Henry Krajeski
}

\maketitle

%===== Justification, spacing for the main text
\raggedbottom
\onehalfspacing
\pagenumbering{roman}

%===== Abstract, Sommaire & Acknowledgments
\section*{\centering Abstract}

This thesis project presents MapperGUI, a cross-platform graphical tool for the manipulation of musical mapping networks. Most digital musical instruments (DMIs) gather gestural input from musicians by way of electronic sensors and transform these data into sound through separate synthesis engines. The mapping of control inputs to synthesis parameters is arbitrary, multi-faceted and extremely important for the effectiveness of DMIs. Software tools exist to aid in this process, they attempt to render the task of musical mapping more transparent, swift and configurable. 

The libmapper software library, developed at the Input Devices and Music Interaction Laboratory, creates a standard framework for DMIs to communicate data on a distributed network and map their signals collaboratively in real-time. MapperGUI presents a graphical user interface for libmapper networks, allowing non-expert users to manipulate the text-based system. The interface aims to be flexible, such that it can accommodate the vast array of musical networks and tasks that must be performed when mapping. To this end, it provides multiple independent visualizations and interaction modes within a single framework. 

This document explores some of the issues challenging the field of musical mapping and describes the motivations behind the MapperGUI project in this context. Relevant research in the fields of data visualization and interface design is summarized and applied to the task of creating a graphical user interface for libmapper networks. Prior graphical interfaces for libmapper are examined for successful features that can be incorporated into MapperGUI. Specific implementation challenges and features of the final program are described. Insight gained from interviews with users of MapperGUI is presented, along with future work and possible extensions for the interface.

MapperGUI is available for free download as a standalone application at \url{www.libmapper.org/downloads}. All code is open-source and can be accessed at \url{https://github.com/mysteryDate/webmapper}.

\newpage

\section*{\centering R\'esum\'e}

Je suis une grande pomme de terre.

%Thesis regulations require that contributions by others in the collection of materials and data, the design and construction of apparatus, the performance of experiments, the analysis of data, and the preparation of the thesis be acknowledged.
\pagebreak

\section*{\centering Acknowledgments}

Many thanks to my thesis advisor, Professor Marcelo M. Wanderley, who directed me towards this project, offered

The work presented here would certainly not exist if not for the thousands of hours spent developing libmapper itself. 

\begin{table}
    \centering
    \begin{tabular}{p{7cm} l}
        \hline\hline
        name&role\\
        \hline
        Marcelo Wanderley& Advisor, IDMIL\\
        Joseph Malloch& libmapper, Maxmapper\\
        Stephen Sinclair& libmapper, Webmapper\\
        Vijay Ruraraju& Vizmapper\\
        Jon Wilansky& GridView, HiveView\\
        H\aa kon Knutzen, Mailis Rodrigues, Clayton Mamedes, Julie Ren\'e&Feedback\\
        Caitlin Stall-Paquet&Translation, editing, being a poulooze\\
    \end{tabular}{l l}
\end{table}

%Thesis regulations require that contributions by others in the collection of materials and data, the design and construction of apparatus, the performance of experiments, the analysis of data, and the preparation of the thesis be acknowledged.
\pagebreak


%========== Tables of contents, figures, tables
\tableofcontents
\listoffigures

\listoftables

\newpage
\chapter*{List of Acronyms}\markright{List of Terms}

\begin{longtable}{ll}
    DMI 	& 	Digital Musical Instrument\\
    GUI		& 	Graphical User Interface\\
    IDMIL &   Input Devices and Music Interaction Laboratory\\ 
    API   &   Application Programming Interface\\
    SWIG  &   Simplified Wrapper and Interface Generator\\
    OSC   &   Open Sound Control\\
    MVC   &   Model View Controller\\
    HTTP  &   HyperText Transfer Protocol\\
    HTML  &   HyperText Markup Language\\
    CSS   &   Cascading Style Sheet\\
    URL   &   Uniform Resource Locater\\
\end{longtable}

\cleardoublepage
\pagenumbering{arabic}

%%========== Chapters
%\typeout{}
%\include{chapters/mikes/C-1}
%
\typeout{}
\include{chapters/introduction}
%
\typeout{}
\include{chapters/background}
%
\typeout{}
\include{chapters/libmapper}
%
\typeout{}
\include{chapters/design}
%
\typeout{}
\include{chapters/results}
%
\typeout{}
\include{chapters/conclusions}
%%==========



%%========== Appendices
%\appendix
%
%%==========
%\typeout{}
%\input{A-A}

%========== Bibliography
\typeout{}
%\begin{singlespace}
    \renewcommand\refname{References}
    \nocite{*}
    \bibliography{../citations/references}%,../citations/other_references}
    \bibliographystyle{Style/ichago}
%\end{singlespace}

\end{document}
